### 2017-08-31 paper.tex
:art: Hydroplaning

\subsection{Water entrainment}
Comparing the snapshots of the granular flow at time $t=3\tau_c$ for the dense column cases~\cref{fig:dense-s5-r09} and the loose column cases~\cref{fig:slope_loose_5,fig:dense-s5-r09}, it can be seen that the overall shapes of the flow front are different.~\Cref{fig:slope-outline} shows the boundary of the flowing mass and its evolution with time for the slope angle \si{5}{\degree} case. It shows that the dense granular columns exhibit a more acute angle of attack (flow front) in contrast to a more parabolic profile for the loose granular columns. This difference in angle of flow front changes the ratio of amount of water entering the flowing granular mass to the water interacting with the surface of the granular mass.~\Cref{fig:water_entrainment_loose_dense} shows the horizontal velocity of the fluid and streamlines at the flow front of the dense and loose granular columns for the slope angle of \si{5}{\degree}. The density of streamlines entering the granular front is significantly higher for the loose column case, in comparison to the dense column case. This clearly shows the effect of the shape of flow front on water entrainment. More water entrainment results in increase in pore space (or decrease in solid fraction) at the flow front, which in turn reduces the interparticle shear resistance.

>>>>

\subsection{Runout phase}
In the second phase, the initial vertical collapse motion is converted to horizontal acceleration. The flow is fully mobilised at $3\tau_c$, when the kinetic energy is at its peak. In the submerged cases, the dynamics of the granular flow are controlled by the following three factors: (a) hydroplaning, defined as the loss of friction between the flowing material and the bottom surface due to the presence of water, (b) water entrainment at the front of the flowing mass, which results in a decrease in the density of the flowing granular mass, and (c) the interaction of the surface of granular flow with the surrounding fluid resulting in formation of eddies and drag effects. The role of water on the dynamics of the dense and loose granular columns is discussed in this section.

\subsubsection{Hydroplaning}
Hydroplaning refers to the loss of friction, which occurs due to the entrapment of water between the granular mass and the slope, which results in zero effective stresses.~\Cref{fig:slope-hydroplaning} shows the effective stress of the flow front for a distance of $~15d$ for the slope angle \si{5}{\degree} case. The effective stress is computed from the contact forces acting between the grains and the bottom surface. The horizontal axis is normalised to the maximum length of $~15d$ from the flow front and the vertical axis is normalised to the initial maximum effective stress. The average effective stress at the bottom of the flowing granular mass over 5 time-steps is presented to avoid fluctuations in the stresses. The loose granular columns have zero average effective stress at the flow front ($~7.5d$ from the front). In contrast, the dense granular columns that have positive effective stress at the sliding interface, which creates frictional forces that work against the flow movement. The loose granular flow tends to entrain more water at the base of the flow front, creating a hydroplaning surface. This results in a longer runout distance. This corroborates the influence of hydroplaning on the loose granular columns which show higher sustained peak kinetic energy in comparison with the dense granular columns.

\begin{figure}[htpb]
\centering
\includegraphics[width=0.9\textwidth]{figs/slope-hydroplaning}
\caption{Effective stress at the flow front ($~15d$) between the dense and loose submerged cases for a slope angle of \si{5}{\degree} at time $t = 3\tau_c$}
\label{fig:slope-hydroplaning}
\end{figure}

\subsubsection{Water entrainment}
Comparing the snapshots of the granular flow at time $t=3\tau_c$ for the dense column cases~\cref{fig:dense-s5-r09} and the loose column cases~\cref{fig:slope_loose_5}, it can be seen that the overall shapes of the flow front are different.~\Cref{fig:slope-outline} shows the boundary of the flowing mass and its evolution with time for the slope angle \si{5}{\degree} case. It shows that the dense granular columns exhibit a more acute angle of attack (flow front) in contrast to a more parabolic profile for the loose granular columns. This difference in angle of flow front changes the ratio of amount of water entering the flowing granular mass to the water interacting with the surface of the granular mass.~\Cref{fig:water_entrainment_loose_dense} shows the horizontal velocity of the fluid and streamlines at the flow front of the dense and loose granular columns for the slope angle of \si{5}{\degree}. The density of streamlines entering the granular front is significantly higher for the loose column case, in comparison to the dense column case. This clearly shows the effect of the shape of flow front on water entrainment. More water entrainment results in increase in pore space (or decrease in solid fraction) at the flow front, which in turn reduces the interparticle shear resistance.
